---
title: "SSL Survey Analysis"
output: html_document
date: "2026-01-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(survey)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(scales)
library(tidyr)
library(marginaleffects)
library(broom)
library(stargazer)

# ==============================================================================
# DEFINE COLORS & THEME
# ==============================================================================

# --- Define Custom Colors First ---
mine_hex_list <- c(
  "#00A896", # Teal
  "#826AED", # Purple
  "#E3B505", # Yellow
  "#89DAFF", # Light Blue
  "#F45B69", # Pink
  "#F18805", # Orange
  "#D41876", # Magenta
  "#020887"  # Dark Blue
)

# --- Define Scale Functions ---
scale_fill_mine <- function(...) {
  scale_fill_manual(values = mine_hex_list, ...)
}

scale_color_mine <- function(...) {
  scale_color_manual(values = mine_hex_list, ...)
}

# --- Define Custom Theme (Fixed) ---
theme_mine <- function(font_scale = 1,
                       font_hierarchy_ratio = 1.25,
                       smallest_font_size = 12,
                       text_color = '#424242',
                       font_family = 'sans', 
                       bold_title = TRUE,
                       save_transparent = FALSE, 
                       figsize = NULL, 
                       base_reference_size = c(9, 6)) {
  
  # Auto-Scaling Logic
  if (!is.null(figsize)) {
    ref_diag <- sqrt(base_reference_size[1]^2 + base_reference_size[2]^2)
    target_diag <- sqrt(figsize[1]^2 + figsize[2]^2)
    scale_factor <- target_diag / ref_diag
    font_scale <- font_scale * scale_factor
  }
  
  # Hierarchy sizes
  small_size <- smallest_font_size * font_scale
  medium_size <- small_size * font_hierarchy_ratio
  large_size  <- medium_size * font_hierarchy_ratio
  
  # Base theme
  theme_minimal(base_size = small_size, base_family = font_family) +
    theme(
      # Text
      text = element_text(color = text_color),
      axis.title = element_text(size = medium_size, color = text_color),
      axis.text = element_text(size = small_size, color = text_color),
      plot.title = element_text(
        size = large_size,
        face = if (bold_title) "bold" else "plain",
        hjust = 0,
        margin = margin(b = 10 * font_scale),
        color = text_color
      ),
      legend.text = element_text(size = small_size),
      legend.title = element_text(size = small_size, face = "bold"),
      
      # Grid & Spines (Fixed: using linewidth)
      panel.grid.major = element_line(
        color = scales::alpha("black", 0.2), 
        linewidth = 0.5 * font_scale, 
        linetype = "dotted"
      ),
      panel.grid.minor = element_blank(),
      panel.border = element_blank(),
      
      # Axis Lines
      axis.line.x.bottom = element_line(color = text_color, linewidth = 0.8 * font_scale),
      axis.line.y.left   = element_line(color = text_color, linewidth = 0.8 * font_scale),
      
      # Backgrounds (Fixed: alpha inside color func, not as argument)
      panel.background = element_rect(fill = "white", color = NA),
      plot.background = element_rect(
        fill = if (save_transparent) "transparent" else "white", 
        color = NA
      ),
      
      # Legend (Fixed: alpha inside color func)
      legend.background = element_rect(fill = scales::alpha("white", 0.95), color = NA),
      legend.key = element_rect(fill = "white", color = NA),
      legend.position = "bottom"
    )
}
```

## Load Data

```{r load-data}
raw_path <- "../data/qualtrics_experiments/[ssl]+ssl_survey+-+ShortForm_January+4,+2026_14.08.csv"

raw_full <- read_csv(raw_path, col_names = TRUE, show_col_types = FALSE)
raw <- raw_full %>% slice(3:n())
```

## Extract Variables

```{r extract-variables}
df <- raw %>%
  select(
    pid,
    duration = `Duration (in seconds)`,
    any_llm = dv_llm_usage,
    moral = dv_use_moral,
    personal = dv_use_personal,
    con = dv_use_con,
    commit,
    attn1
  )
```

## Create Usage Variables

```{r usage-variables}
# Map for general LLM usage frequency
LLM_FREQ_MAP <- c(
  "Yes - almost constantly" = 5,
  "Yes - several times a day" = 4,
  "Yes - about once a day" = 3,
  "Yes - several times a week" = 2,
  "Yes - less often than several times a week" = 1,
  "No - never used" = 0
)

# Map for domain-specific usage counts
DOMAIN_USAGE_MAP <- c(
  "Yes - once or twice" = 1,
  "Yes - 3-10 times" = 2,
  "Yes - 11-20 times" = 3,
  "Yes - more than 20 times" = 4,
  "No - never" = 0
)

df <- df %>%
  mutate(
    duration = as.numeric(duration) / 60,
    any_llm_num = LLM_FREQ_MAP[any_llm],
    moral_num = DOMAIN_USAGE_MAP[moral],
    personal_num = DOMAIN_USAGE_MAP[personal],
    con_num = DOMAIN_USAGE_MAP[con],
    any_any_llm = if_else(any_llm_num > 0, 1, 0),
    any_moral = if_else(moral_num > 0, 1, 0),
    any_personal = if_else(personal_num > 0, 1, 0),
    any_con = if_else(con_num > 0, 1, 0),
    any_usage = pmax(any_moral, any_personal, any_con, na.rm = TRUE)
  )
```

## Apply Filters

```{r apply-filters}
df <- df %>%
  filter(attn1 == '25' & commit == 'Yes, I will')

cat("Filtering out non-llm users\n")
cat(paste("Original N:", nrow(df), "\n"))

df <- df %>%
  filter(any_llm != 'No - never used')

cat(paste("N after filtering:", nrow(df), "\n\n"))
```

## Compute Weights Using Raking

```{r compute-weights-raking}
```{r compute-weights-raking}
library(survey)

# 1. Define Raw Population Ratios (Relative importance/frequency)
# These do not need to sum to 1 or N yet; we will scale them.
raw_pop_ratios <- c(
  "Yes - almost constantly" = 5,
  "Yes - several times a day" = 7,
  "Yes - about once a day" = 5,
  "Yes - several times a week" = 10,
  "Yes - less often than several times a week" = 25
)

# 2. Align Population Targets with Sample Data
# We filter the targets to ensure we only target categories that actually exist 
# in the current filtered dataframe (preventing 0-division errors).
sample_levels <- unique(as.character(df$any_llm))
matched_ratios <- raw_pop_ratios[names(raw_pop_ratios) %in% sample_levels]

# 3. Create the Population Dataframe (The "Control" Total)
# CRITICAL: We scale the frequencies so they sum to the current sample size (nrow(df)).
# This ensures the resulting weights average to 1.
target_n <- nrow(df) 
scaled_freqs <- (matched_ratios / sum(matched_ratios)) * target_n

pop_llm <- data.frame(
  any_llm = names(scaled_freqs),
  Freq = as.numeric(scaled_freqs)
)

print("Target Frequencies (Sum should match Sample N):")
print(pop_llm)

# 4. Create Unweighted Design
# Ensure the variable in df is a factor with the same levels as the target
df$any_llm <- factor(df$any_llm, levels = pop_llm$any_llm)
d_unw <- svydesign(ids = ~1, data = df)

# 5. Apply Raking
d_rake <- rake(
  design = d_unw,
  sample.margins = list(~any_llm),
  population.margins = list(pop_llm),
  control = list(maxit = 50, epsilon = 1e-7)
)

# 6. Validate Weights
# The sum of weights should equal the sample size
cat(paste("\nSum of weights:", round(sum(weights(d_rake)), 2)))
cat(paste("\nSample size:   ", nrow(df)))
cat("\n\nWeight Summary:\n")
print(summary(weights(d_rake)))
```

## Usage Rates

```{r usage-rates}
for (use_type in c('moral', 'personal', 'con')) {
  cat(paste0("\n", toupper(use_type), " USAGE:\n"))
  
  # Unweighted
  unweighted_tab <- prop.table(table(df[[use_type]])) * 100
  cat("\nUnweighted:\n")
  print(round(unweighted_tab, 2))
  
  # Weighted using survey package
  weighted_result <- svytable(as.formula(paste0("~", use_type)), svy_design)
  weighted_pct <- prop.table(weighted_result) * 100
  cat("\nWeighted:\n")
  print(round(weighted_pct, 2))
  cat("\n")
}
```

## Any Usage Summary

```{r any-usage}
cat("\n=== Any Usage (Moral OR Personal OR Con) ===\n\n")

# Unweighted
# make df2 which is 6 times df
unweighted_mean <- mean(df$any_usage)
unweighted_se <- sqrt(unweighted_mean * (1 - unweighted_mean) / nrow(df))
unweighted_moe <- 1.96 * unweighted_se

# Weighted using survey package
weighted_result <- svymean(~any_usage, svy_design)
weighted_mean <- as.numeric(weighted_result)
weighted_se <- as.numeric(SE(weighted_result))
weighted_moe <- 1.96 * weighted_se

# Create summary table
summary_table <- data.frame(
  Type = c("Unweighted", "Weighted"),
  Mean = round(c(unweighted_mean, weighted_mean) * 100, 2),
  SE = round(c(unweighted_se, weighted_se) * 100, 2),
  MoE = round(c(unweighted_moe, weighted_moe) * 100, 2),
  CI_Lower = round(c(unweighted_mean - unweighted_moe, 
                     weighted_mean - weighted_moe) * 100, 2),
  CI_Upper = round(c(unweighted_mean + unweighted_moe, 
                     weighted_mean + weighted_moe) * 100, 2)
)

colnames(summary_table) <- c("Type", "Mean", "SE", "MoE", "CI_Lower", "CI_Upper")

# make ggplot graph
p <- ggplot(summary_table, aes(x = Type, y = Mean, fill = Type)) +
  geom_bar(stat = "identity", position = position_dodge(), width = 0.5) +
  geom_errorbar(aes(ymin = CI_Lower, ymax = CI_Upper), 
                width = 0.2, position = position_dodge(0.5)) +
  labs(title = "Any SSL Usage Rates",
       y = "Mean",
       x = "") +
  theme_mine() +
  scale_fill_mine() +
  theme(legend.position = "none")
print(p)
```



## Standardize Variables and Fit Model

```{r fit-model}
library(marginaleffects)

# Define variable groups
cat_vars <- c("any_llm")
numeric_vars <- c("duration")

# Standardize numeric variables
for (var in numeric_vars) {
  df[[paste0(var, "_z")]] <- scale(df[[var]])[,1]
}

# Update survey design with standardized variables
svy_design <- rake(
  design = svydesign(ids = ~1, data = df),
  sample.margins = list(~any_llm),
  population.margins = list(pop_llm),
  control = list(maxit = 50, epsilon = 1e-7)
)

# Fit logistic model

formula_obj <- as.formula("any_usage ~ any_llm + duration_z")
logistic_model.survey <- svyglm(formula_obj, design = svy_design, family = binomial())
logistic_model.raw <- glm(formula_obj, data = df, family = binomial())

cat("\n=== Model Summary ===\n")
print(summary(logistic_model))

library(stargazer)
s <- stargazer(logistic_model.raw, logistic_model.survey,
          type = "text",
          title = "Logistic Regression Results",
          dep.var.labels = c("Any SSL Usage"),
          covariate.labels = c("Any LLM Usage", "Duration (standardized)"),
          out = NULL)
print(s)
```

## Average Marginal Effects

```{r ame-plot, fig.width=10, fig.height=6}
library(ggplot2)

# Compute AMEs for all variables
ame_results <- avg_comparisons(logistic_model)

# Extract results as dataframe
ame_df <- as.data.frame(ame_results)

cat("\n=== Average Marginal Effects ===\n")
print(ame_df[, c("term", "contrast", "estimate")])

####################
# Prepare Plot Data
####################

# Add variable type
ame_df$var_type <- ifelse(
  gsub("_z$", "", ame_df$term) %in% cat_vars,
  "categorical",
  "numeric"
)

# Create variable grouping (clean names)
ame_df$variable <- gsub("_z$", "", ame_df$term)

# Create simple labels with variable name embedded
ame_df$display_label <- NA

for (i in 1:nrow(ame_df)) {
  var_name <- ame_df$term[i]
  
  if (ame_df$var_type[i] == "categorical") {
    contrast_text <- as.character(ame_df$contrast[i])
    parts <- strsplit(contrast_text, " - ")[[1]]
    
    if (length(parts) >= 2) {
      midpoint <- ceiling(length(parts) / 2)
      comparison <- paste(parts[1:midpoint], collapse = " - ")
      reference <- paste(parts[(midpoint+1):length(parts)], collapse = " - ")
      
      ame_df$display_label[i] <- paste0(
        var_name, ": ", comparison, "\n(ref: ", reference, ")"
      )
    } else {
      ame_df$display_label[i] <- paste0(var_name, ": ", contrast_text)
    }
    
  } else {
    # For numeric
    var_name_clean <- gsub("_z$", "", var_name)
    ame_df$display_label[i] <- paste0(var_name_clean, ": +1 SD")
  }
}

# Order by variable type and variable
ame_df <- ame_df %>%
  arrange(desc(var_type), variable) %>%
  mutate(row_id = row_number())

####################
# Plot AMEs
####################

# Define colors for each variable
var_colors <- c(
  "any_llm" = mine_hex_list[1],
  "duration" = mine_hex_list[2]
)

p <- ggplot(ame_df, aes(x = estimate, y = reorder(display_label, row_id))) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_errorbarh(
    aes(xmin = conf.low, xmax = conf.high, color = variable),
    height = 0,
    linewidth = 1
  ) +
  geom_point(
    aes(color = variable),
    size = 3
  ) +
  scale_color_manual(
    values = var_colors,
    name = "Variable"
  ) +
  labs(
    title = "Average Marginal Effects on SSL Usage",
    x = "Average Marginal Effect (Percentage Points)",
    y = NULL,
    caption = "Note: Numeric variables standardized (M=0, SD=1)"
  ) +
  theme_mine() +
  theme(
    legend.position = "bottom",
    panel.grid.major.y = element_blank(),
    axis.text.y = element_text(hjust = 1, size = 10)
  )

print(p)
```