---
title: "SSL Survey Analysis"
output: 
  html_document: 
    toc: true
date: "2026-01-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(survey)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(scales)
library(tidyr)
library(marginaleffects)
library(broom)
library(stargazer)
library(readr)

# ==============================================================================
# DEFINE COLORS & THEME
# ==============================================================================

# --- Define Custom Colors First ---
mine_hex_list <- c(
  "#00A896", # Teal
  "#826AED", # Purple
  "#E3B505", # Yellow
  "#89DAFF", # Light Blue
  "#F45B69", # Pink
  "#F18805", # Orange
  "#D41876", # Magenta
  "#020887"  # Dark Blue
)

# --- Define Scale Functions ---
scale_fill_mine <- function(...) {
  scale_fill_manual(values = mine_hex_list, ...)
}

scale_color_mine <- function(...) {
  scale_color_manual(values = mine_hex_list, ...)
}

# --- Define Custom Theme (Fixed) ---
theme_mine <- function(font_scale = 1,
                       font_hierarchy_ratio = 1.25,
                       smallest_font_size = 12,
                       text_color = '#424242',
                       font_family = 'sans', 
                       bold_title = TRUE,
                       save_transparent = FALSE, 
                       figsize = NULL, 
                       base_reference_size = c(9, 6)) {
  
  # Auto-Scaling Logic
  if (!is.null(figsize)) {
    ref_diag <- sqrt(base_reference_size[1]^2 + base_reference_size[2]^2)
    target_diag <- sqrt(figsize[1]^2 + figsize[2]^2)
    scale_factor <- target_diag / ref_diag
    font_scale <- font_scale * scale_factor
  }
  
  # Hierarchy sizes
  small_size <- smallest_font_size * font_scale
  medium_size <- small_size * font_hierarchy_ratio
  large_size  <- medium_size * font_hierarchy_ratio
  
  # Base theme
  theme_minimal(base_size = small_size, base_family = font_family) +
    theme(
      # Text
      text = element_text(color = text_color),
      axis.title = element_text(size = medium_size, color = text_color),
      axis.text = element_text(size = small_size, color = text_color),
      plot.title = element_text(
        size = large_size,
        face = if (bold_title) "bold" else "plain",
        hjust = 0,
        margin = margin(b = 10 * font_scale),
        color = text_color
      ),
      legend.text = element_text(size = small_size),
      legend.title = element_text(size = small_size, face = "bold"),
      
      # Grid & Spines (Fixed: using linewidth)
      panel.grid.major = element_line(
        color = scales::alpha("black", 0.2), 
        linewidth = 0.5 * font_scale, 
        linetype = "dotted"
      ),
      panel.grid.minor = element_blank(),
      panel.border = element_blank(),
      
      # Axis Lines
      axis.line.x.bottom = element_line(color = text_color, linewidth = 0.8 * font_scale),
      axis.line.y.left   = element_line(color = text_color, linewidth = 0.8 * font_scale),
      
      # Backgrounds (Fixed: alpha inside color func, not as argument)
      panel.background = element_rect(fill = "white", color = NA),
      plot.background = element_rect(
        fill = if (save_transparent) "transparent" else "white", 
        color = NA
      ),
      
      # Legend (Fixed: alpha inside color func)
      legend.background = element_rect(fill = scales::alpha("white", 0.95), color = NA),
      legend.key = element_rect(fill = "white", color = NA),
      legend.position = "bottom"
    )
}
```

## Load Data

```{r load-data}
raw_path <- "../data/qualtrics_experiments/[ssl]+ssl_survey+-+ShortForm_January+4,+2026_15.49.csv"

raw_full <- read_csv(raw_path, col_names = TRUE, show_col_types = FALSE)
raw <- raw_full %>% slice(3:n())

df <- raw %>%
  select(
    pid,
    duration = `Duration (in seconds)`,
    any_llm = dv_llm_usage,
    moral = dv_use_moral,
    personal = dv_use_personal,
    con = dv_use_con,
    commit,
    attn1
  )
```


## Create Usage Variables

```{r usage-variables}
# Map for general LLM usage frequency
LLM_FREQ_MAP <- c(
  "Yes - almost constantly" = 5,
  "Yes - several times a day" = 4,
  "Yes - about once a day" = 3,
  "Yes - several times a week" = 2,
  "Yes - less often than several times a week" = 1,
  "No - never used" = 0
)

# Map for domain-specific usage counts
DOMAIN_USAGE_MAP <- c(
  "Yes - once or twice" = 1,
  "Yes - 3-10 times" = 2,
  "Yes - 11-20 times" = 3,
  "Yes - more than 20 times" = 4,
  "No - never" = 0
)

df <- df %>%
  mutate(
    duration = as.numeric(duration) / 60,
    any_llm_num = LLM_FREQ_MAP[any_llm],
    moral_num = DOMAIN_USAGE_MAP[moral],
    personal_num = DOMAIN_USAGE_MAP[personal],
    con_num = DOMAIN_USAGE_MAP[con],
    any_any_llm = if_else(any_llm_num > 0, 1, 0),
    any_moral = if_else(moral_num > 0, 1, 0),
    any_personal = if_else(personal_num > 0, 1, 0),
    any_con = if_else(con_num > 0, 1, 0),
    any_usage = pmax(any_moral, any_personal, any_con, na.rm = TRUE)
  )
```

## Apply Filters

```{r apply-filters}
df <- df %>%
  filter(attn1 == '25' & commit == 'Yes, I will')

cat("Filtering out non-llm users\n")
cat(paste("Original N:", nrow(df), "\n"))

df <- df %>%
  filter(any_llm != 'No - never used')

cat(paste("N after filtering:", nrow(df), "\n\n"))
```

## Compute Weights Using Raking

```{r compute-weights-raking}
library(survey)

# Define population PROPORTIONS (not counts)
LLM_USAGE_PROPS <- c(
  "Yes - almost constantly" = 5,
  "Yes - several times a day" = 7,
  "Yes - about once a day" = 5,
  "Yes - several times a week" = 10,
  "Yes - less often than several times a week" = 25
)

# Rescale to sum to sample size
LLM_USAGE_COUNTS <- (LLM_USAGE_PROPS / sum(LLM_USAGE_PROPS)) * nrow(df)

# Filter to sample categories
sample_categories <- names(table(df$any_llm))
LLM_USAGE_COUNTS <- LLM_USAGE_COUNTS[names(LLM_USAGE_COUNTS) %in% sample_categories]

# Convert to data frame
pop_llm <- data.frame(
  any_llm = names(LLM_USAGE_COUNTS),
  Freq = LLM_USAGE_COUNTS
)

cat("\n=== Population Margins Data Frame ===\n")
print(pop_llm)

####################
# Create Unweighted Survey Design
####################

svy_unweighted <- svydesign(
  ids = ~1,       
  data = df
)

cat("\n=== Unweighted Design Summary ===\n")
cat("N:", nrow(df), "\n")
cat("All weights = 1 (unweighted)\n")

####################
# Apply Raking (Post-Stratification Weighting)
####################

cat("\n=== Applying Raking Weights ===\n")

svy_design <- rake(
  design = svy_unweighted,
  sample.margins = list(~any_llm),          # Variable to weight on
  population.margins = list(pop_llm),       # Target distribution
  control = list(maxit = 50, epsilon = 1e-7)  # Convergence parameters
)

cat("Raking converged successfully.\n")

####################
# Check Resulting Weights
####################

cat("\n=== Weight Summary ===\n")
print(summary(weights(svy_design)))

cat("\n=== Weighted Distribution (After Raking) ===\n")
weighted_dist <- svytable(~any_llm, svy_design)
cat("Weighted frequencies:\n")
print(weighted_dist)
cat("\nWeighted proportions:\n")
print(prop.table(weighted_dist) * 100)

cat("\n=== Comparison: Target vs Weighted ===\n")
comparison <- data.frame(
  Category = names(LLM_USAGE_COUNTS),
  Target_Prop = round(prop.table(LLM_USAGE_COUNTS) * 100, 2),
  Weighted_Prop = round(prop.table(weighted_dist) * 100, 2)
)
print(comparison)
```

## Usage Rates

```{r usage-rates}
for (use_type in c('moral', 'personal', 'con')) {
  cat(paste0("\n", toupper(use_type), " USAGE:\n"))
  
  # Unweighted
  unweighted_tab <- prop.table(table(df[[use_type]])) * 100
  cat("\nUnweighted:\n")
  print(round(unweighted_tab, 2))
  
  # Weighted using survey package
  weighted_result <- svytable(as.formula(paste0("~", use_type)), svy_design)
  weighted_pct <- prop.table(weighted_result) * 100
  cat("\nWeighted:\n")
  print(round(weighted_pct, 2))
  cat("\n")
}
```

## Any Usage Summary

```{r any-usage}
cat("\n=== Any Usage (Moral OR Personal OR Con) ===\n\n")

# Unweighted
# make df2 which is 6 times df
unweighted_mean <- mean(df$any_usage)
unweighted_se <- sqrt(unweighted_mean * (1 - unweighted_mean) / nrow(df))
unweighted_moe <- 1.96 * unweighted_se

# Weighted using survey package
weighted_result <- svymean(~any_usage, svy_design)
weighted_mean <- as.numeric(weighted_result)
weighted_se <- as.numeric(SE(weighted_result))
weighted_moe <- 1.96 * weighted_se

# Create summary table
summary_table <- data.frame(
  Type = c("Unweighted", "Weighted"),
  Mean = round(c(unweighted_mean, weighted_mean) * 100, 2),
  SE = round(c(unweighted_se, weighted_se) * 100, 2),
  MoE = round(c(unweighted_moe, weighted_moe) * 100, 2),
  CI_Lower = round(c(unweighted_mean - unweighted_moe, 
                     weighted_mean - weighted_moe) * 100, 2),
  CI_Upper = round(c(unweighted_mean + unweighted_moe, 
                     weighted_mean + weighted_moe) * 100, 2)
)

colnames(summary_table) <- c("Type", "Mean", "SE", "MoE", "CI_Lower", "CI_Upper")

# make ggplot graph
p <- ggplot(summary_table, aes(x = Type, y = Mean, fill = Type)) +
  geom_bar(stat = "identity", position = position_dodge(), width = 0.5) +
  geom_errorbar(aes(ymin = CI_Lower, ymax = CI_Upper), 
                width = 0.2, position = position_dodge(0.5)) +
  labs(title = "Any SSL Usage Rates",
       y = "Mean",
       x = "") +
  theme_mine() +
  scale_fill_mine() +
  theme(legend.position = "none") + 
  # add exact number to graph
  geom_text(aes(label = paste0(Mean, "%")), 
            vjust = 6.5, 
            color = "white",
            position = position_dodge(0.5), 
            size = 5)

print(p)
print(summary_table)
```



## Standardize Variables and Fit Model

```{r fit-model}

# Define variable groups
cat_vars <- c("any_llm")
numeric_vars <- c("duration")

llm_order <- c(
  "Yes - less often than several times a week",
  "Yes - several times a week",
  "Yes - about once a day",
  "Yes - several times a day",
  "Yes - almost constantly"
)
# Convert to ordered factor
df$any_llm <- factor(df$any_llm, levels = llm_order, ordered = FALSE)


# Standardize numeric variables
for (var in numeric_vars) {
  df[[paste0(var, "_z")]] <- scale(df[[var]])[,1]
}

# Update survey design with standardized variables
svy_design <- rake(
  design = svydesign(ids = ~1, data = df),
  sample.margins = list(~any_llm),
  population.margins = list(pop_llm),
  control = list(maxit = 50, epsilon = 1e-7)
)

# Fit logistic model

formula_obj <- as.formula("any_usage ~ any_llm + duration_z")
logistic_model.survey <- svyglm(formula_obj, design = svy_design, family = binomial())
logistic_model.raw <- glm(formula_obj, data = df, family = binomial())


s <- stargazer(logistic_model.raw, logistic_model.survey,
          type = "text",
          title = "Logistic Regression Results",
          dep.var.labels = c("Any SSL Usage"),
          out = NULL)
print(s)
```

## Average Marginal Effects

```{r ame-plot, fig.width=10, fig.height=6}
library(ggplot2)

# Compute AMEs for all variables
ame_results <- avg_comparisons(logistic_model.survey )

# Extract results as dataframe
ame_df <- as.data.frame(ame_results)

cat("\n=== Average Marginal Effects ===\n")
print(ame_df[, c("term", "contrast", "estimate")])

####################
# Prepare Plot Data
####################

# Add variable type
ame_df$var_type <- ifelse(
  gsub("_z$", "", ame_df$term) %in% cat_vars,
  "categorical",
  "numeric"
)

# Create variable grouping (clean names)
ame_df$variable <- gsub("_z$", "", ame_df$term)

# Create simple labels with variable name embedded
ame_df$display_label <- NA

for (i in 1:nrow(ame_df)) {
  var_name <- ame_df$term[i]
  
  if (ame_df$var_type[i] == "categorical") {
    contrast_text <- as.character(ame_df$contrast[i])
    parts <- strsplit(contrast_text, " - ")[[1]]
    
    if (length(parts) >= 2) {
      midpoint <- ceiling(length(parts) / 2)
      comparison <- paste(parts[1:midpoint], collapse = " - ")
      reference <- paste(parts[(midpoint+1):length(parts)], collapse = " - ")
      
      ame_df$display_label[i] <- paste0(
        var_name, ": ", comparison, "\n(ref: ", reference, ")"
      )
    } else {
      ame_df$display_label[i] <- paste0(var_name, ": ", contrast_text)
    }
    
  } else {
    # For numeric
    var_name_clean <- gsub("_z$", "", var_name)
    ame_df$display_label[i] <- paste0(var_name_clean, ": +1 SD")
  }
}

# Order by variable type and variable
ame_df <- ame_df %>%
  arrange(desc(var_type), variable) %>%
  mutate(row_id = row_number())

####################
# Plot AMEs
####################

# Define colors for each variable
var_colors <- c(
  "any_llm" = mine_hex_list[1],
  "duration" = mine_hex_list[2]
)

p <- ggplot(ame_df, aes(x = estimate, y = reorder(display_label, row_id))) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_errorbarh(
    aes(xmin = conf.low, xmax = conf.high, color = variable),
    height = 0,
    linewidth = 1
  ) +
  geom_point(
    aes(color = variable),
    size = 3
  ) +
  scale_color_manual(
    values = var_colors,
    name = "Variable"
  ) +
  labs(
    title = "Average Marginal Effects on SSL Usage",
    x = "Average Marginal Effect (Percentage Points)",
    y = NULL,
    caption = "Note: Numeric variables standardized (M=0, SD=1)"
  ) +
  theme_mine() +
  theme(
    legend.position = "bottom",
    panel.grid.major.y = element_blank(),
    axis.text.y = element_text(hjust = 1, size = 10)
  )

print(p)
```


# Simulation of total conversations


The total number of SSL conversations is estimated as the product of three things, (A) x (B) x (C):
- (A) The number of U.S. adults (approximately 260 million)
- (B) The proportion of U.S. adults who engage in SSL (from prior study)
- (C) The average number of SSL conversations per SSL user (from current survey).

We use Monte Carlo simulations to propagate uncertainty.  For (B), we sample from a normal distribution parameterized by the mean and standard error from a prior study. For (C),  respondents selected categorical frequency ranges (e.g., '3-10 times'). We sample uniformly within each respondent's reported range. Across 10k iterations, we then multiply (A) x (B) x (C) to arrive at a distribution of total SSL conversations.

We make one major conservative assumption that makes our estimate likely a lower bound. When a respondent reported different frequencies across domains (moral, personal, conventional), we take the maximum rather than the sum, assuming complete overlap. For example, if a participant said "Between 3-10 times" for moral, personal, conventional, we go with "3-10" rather than sum({3-10, 3-10, 3-10}). Intuitively, this is summing that a person had 3-10 total SSL conversations that overlapped across each domain.


### Algorithm

Input:
- `A`: Number of U.S. adults (constant = 260,000,000)
- `mu_B, sigma_B`: Mean and SE of SSL prevalence from prior study
- `S`: Survey data with `n` respondents, each reporting frequency category
- `R`: Mapping from categories to numeric ranges
- `N`: Number of Monte Carlo iterations (= 10,000)

Procedure:

1. For each respondent `i` in `S`:
   ```
   range[i] <- maximum frequency range (moral, personal, conventional)
   ```

2. For iteration `j = 1` to `N`:
   
   (a) For each respondent `i` in `S`:
   ```
   convos[i] ← uniform(range.lower, range.upper)
   ```
   
   (b) Compute average conversations per SSL user:
   ```
   C_j ← mean(convos)
   ```
   
   (c) Sample prevalence (truncated to [0,1]):
   ```
   B_j <-  Normal(u_B, sigma_B) clipped to 1
   ```
   
   (d) Compute total SSL conversations:
   ```
   T_j <- A × B_j × C_j
   ```

3. Return: Distribution of `{T_1, T_2, ..., T_N}`




```{r}
####################
# Simulation: Total SSL Conversations in US
####################

# Define conversation ranges for each frequency level
CONV_RANGES <- list(
  "Yes - once or twice" = c(1, 2),
  "Yes - 3-10 times" = c(3, 10),
  "Yes - 11-20 times" = c(11, 20),
  "Yes - more than 20 times" = c(21, 25),  # Upper bound assumption
  "No - never" = c(0, 0)
)

# Get max frequency for each person (moral, personal, con)
# (Well, we could sum them, but max is more conservative since we are assuming total overlap)
df$max_domain <- pmax(df$moral_num, df$personal_num, df$con_num, na.rm = TRUE)

# Map back to category labels
df$max_domain_label <- names(DOMAIN_USAGE_MAP)[match(df$max_domain, DOMAIN_USAGE_MAP)]

# US parameters
US_ADULTS <- 260000000  # ~260 million adults

# Prevalence from prior study (sample proportion with sampling uncertainty)
SSL_PREVALENCE_MEAN <- 0.50    # Point estimate
SSL_PREVALENCE_LOWER <- 0.48   # 95% CI lower bound
SSL_PREVALENCE_UPPER <- 0.52   # 95% CI upper bound

# Calculate SE from 95% CI
# CI = mean ± 1.96*SE  =>  SE = (upper - lower) / (2 * 1.96)
SSL_PREVALENCE_SE <- (SSL_PREVALENCE_UPPER - SSL_PREVALENCE_LOWER) / (2 * 1.96)

cat("\n=== Prevalence Parameters ===\n")
cat(sprintf("Mean: %.3f\n", SSL_PREVALENCE_MEAN))
cat(sprintf("95%% CI: [%.3f, %.3f]\n", SSL_PREVALENCE_LOWER, SSL_PREVALENCE_UPPER))
cat(sprintf("SE: %.4f\n", SSL_PREVALENCE_SE))

# Run simulation
n_sims <- 10000
total_convos <- numeric(n_sims)

set.seed(123)  # For reproducibility

for (i in 1:n_sims) {
  # For each person, sample number of conversations uniformly from their range
  person_convos <- sapply(df$max_domain_label, function(label) {
    range <- CONV_RANGES[[label]]
    runif(1, min = range[1], max = range[2])
  })
  
  # Average conversations per SSL user
  avg_convos_per_user <- mean(person_convos)
  
  # Sample prevalence from normal distribution (sampling distribution)
  SSL_PREVALENCE <- rnorm(1, mean = SSL_PREVALENCE_MEAN, sd = SSL_PREVALENCE_SE)
  SSL_PREVALENCE <- max(0, min(1, SSL_PREVALENCE))  # Truncate to [0, 1]
  
  # Total US conversations
  total_convos[i] <- avg_convos_per_user * SSL_PREVALENCE * US_ADULTS
}

# Summary statistics
cat("\nRotal SSL Conversations in US (Simulation)\n")
cat(sprintf("Mean: %.2f billion\n", mean(total_convos) / 1e9))
cat(sprintf("Median: %.2f billion\n", median(total_convos) / 1e9))
cat(sprintf("95%% CI: [%.2f billion, %.2f billion]\n", 
            quantile(total_convos, 0.025) / 1e9,
            quantile(total_convos, 0.975) / 1e9))

# Histogram
subtitle_str = "Based on %d Monte Carlo estimations | Mean: %.0f Million, 95%% CI = [%.0f Million, %.0f Million]"
subtitle_str = sprintf(subtitle_str, n_sims, mean(total_convos) / 1e6,
                       quantile(total_convos, 0.025) / 1e6,
                       quantile(total_convos, 0.975) / 1e6)

p <- ggplot(data.frame(total = total_convos / 1e6), aes(x = total)) +
  geom_histogram(bins = 30, fill = "#377EB8", color = "white") +
  geom_vline(xintercept = mean(total_convos) / 1e6, 
             linetype = "dashed", color = "red", linewidth = 1) +
  labs(
    title = "Total Synthetic Social Learning Conversations in US",
    subtitle = subtitle_str,
    x = "Total Conversations (Millions)",
    y = "Count"
  ) +
  theme_minimal()

print(p)
```

# Does ANY SSL differ by USage intensity? Probably

```{r}

cat("\n=== ANY SSL Usage by LLM Usage Intensity ===\n\n")

usage_levels <- levels(df$any_llm)

ug <- ggplot(
  df,
  aes(x = any_llm, fill = factor(any_usage))
) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = percent_format()) +
  scale_fill_manual(
    name = "Any SSL Usage",
    labels = c("No", "Yes"),
    values = c("#E41A1C", "#377EB8")  # Adjust colors as needed
  ) +
  labs(
    title = "Any SSL Usage by LLM Usage Intensity",
    x = "LLM Usage Frequency",
    y = "Proportion"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

print(ug)

```

